Marginally-Stable Thermal Equilibria of Rayleigh-Bénard Convection

### 2021-04-30 publication_materials/ms_rbc.tex

\author{Liam O'Connor$^1$}
\author{Daniel Lecoanet$^{1, 2}$}
\author{Evan Anders$^2$}
\affiliation{%
$^1$Department of Engineering Sciences and Applied Mathematics, Northwestern University, Evanston, IL 60208 USA}
\affiliation{%
$^2$Center for Interdisciplinary Exploration and Research in Astrophysics, Northwestern University, Evanston, IL, 60201 USA}

\begin{abstract}
    Natural convection is ubiquitous throughout the physical sciences and engineering, yet many of its important properties remain elusive---particularly in the large Rayleigh number regime.
    In this investigation, we derive and solve a quasilinear form of the Rayleigh-Benard problem by representing the perturbations in terms of marginally stable eigenmodes.
    The amplitude of each eigenmode is determined by requiring that the background state maintains marginal stability.
    The background temperature profile evolves due to the advective flux of each eigenmode, as well as diffusion.
    The entire calculation is one-dimensional, and can be run on a workstation.
    We find the background temperature field evolves to an equilibrium state, where the advective flux from the marginally-stable eigenmodes and the diffusive flux sum to a constant.
    These marginally-stable thermal equilibria are exact solutions of the quasilinear equations.
    The mean temperature profile has thinner boundary layers and larger Nusselt numbers than thermally-equilibrated 2D and 3D simulations of the full nonlinear equations.
    We find the Nusselt number scales like $\Nu \sim\Ra^{1/3}$.
    When initializing a 2D simulation with a marginally-stable thermal equilibrium, we find the simulation does not exhibit the initial burst of turbulence typical of simulations initialized with the conductive background state plus noise.
    Although the $\Nu$ quickly equilibrates in these simulations, the kinetic energy evolves on a viscous timescale as domain-filling flywheel flows viscously attenuate.
\end{abstract}

@@

\section{Introduction}
Rayleigh-B\'enard convection plays a foundational role in astrophysical and geophysical settings.
The resulting buoyancy-driven flows regulate heat transfer and generate large-scale vortices \cite{Couston}.
Turbulent convection, which is associated with large Rayleigh numbers $\Ra$, is difficult to simulate. 
State of the art simulations performed by \cite{Zhu_2018} have reached $\Ra \sim 10^{14}$ but estimates for the sun's convective zone and earth's interior are $Ra \sim 10^{16}-10^{20}$ and $Ra \sim 10^{20}-10^{30}$ respectively \cite{Ossendrijver,Gubbins_2001}. 
The scaling behavior of the Nusselt number $\Nu \sim Ra^{\beta}$ in the asymptotic ultimate regime is of particular interest.
There exists a substantial body of work pertaining to this specific topic with no general consensus \cite{Malkus_1954, Howard_1966, Kraichnan, Spiegel, Castaing, Grossman, Ahlers}.

@@

\section{Model Setup}\label{sec:model}
We begin with the non-dimensionalized Boussinesq approximation for Rayleigh-Bénard Convection. 
The domain $\mathcal{D}$ is 2-dimensional, rectangular, and horizontally periodic with spatial dimensions $0 < x < 4$ and $-1/2 < z < 1/2$. 
The fluid of interest is constrained between two flat boundaries at $z = -1/2$ and $z = 1/2$ with fixed temperatures $1/2$ and $-1/2$ respectively. 
At both boundaries we specify impenetrable, no-slip conditions, such that the velocity $\vec{u} = u \hat{x} + w \hat{z} = \vec{0}$ at $z = \pm 1/2$, where $\hat{x}, \hat{z}$ are the unit vectors in the $x$ and $z$ directions. 
The equations of motion are then given by
\begin{align}
    \nabla \cdot \vec{u} &= 0 \label{EQ:motion1}\\
    \frac{\partial \vec{u}}{\partial t} + \vec{u} \cdot \nabla \vec{u} &= - \nabla p + T \hat{z} + \mathcal{R} \nabla^2 \vec{u} \label{EQ:motion2}\\
    \frac{\partial T}{\partial t} + \vec{u} \cdot \nabla T &= \mathcal{P} \nabla^2 T \label{EQ:motion3}
\end{align}
where $p$ is pressure and $T$ is temperature. 
For completeness, we specify a final boundary condition $p = p_0$ at $z = \pm 1/2$. 
Any system of this form can be characterized by its dimensionless Rayleigh number $\Ra = \frac{g\alpha L^3 \Delta T}{\nu \kappa}$ and Prandtl number $\Pr = \frac{\nu}{\kappa}$, where $g, \, \alpha, \, L, \, \Delta T, \nu, \kappa$ are the gravitational acceleration, coefficient of thermal expansion, domain height, opposed temperature difference, kinematic viscosity, and thermal diffusivity respectively. 
In this paper, we fix $\Pr = 1$.
For convenience, we define
\begin{equation}
\mathcal{R} = \sqrt{\frac{\Pr}{\Ra}}, \qquad \mathcal{P} = \frac{1}{\sqrt{\Pr \Ra}}.
\end{equation}

@@

To solve \eq{EQ:T0_IVP} numerically, we need an expression for the perturbation so we can calculate the advective heat flux.
Here we will make the quasilinear approximation, dropping the $\vec{u}'\vec{\cdot}\vec{\nabla}\vec{u}'$ and $\vec{u}'\vec{\cdot}\vec{\nabla}T'$ terms from the evolution equations for the perturbations.
Substituting \eqss{EQ:reynolds_dc_u}{EQ:reynolds_dc_p} into \eqss{EQ:motion1}{EQ:motion3} and subtracting \eq{EQ:T0_IVP} from the resulting temperature equation, and linearizing gives
\begin{align}
    \nabla \cdot \vec{u'} &= 0 \label{EQ:linear1}\\
    \frac{\partial\vec{u'}}{\partial t} &= - \nabla p' + T'\hat{z} + \mathcal{R} \nabla^2 \vec{u'} \label{EQ:linear2}\\
    \frac{\partial T'}{\partial t} + \frac{\partial \bar{T}}{\partial z} w' &= \mathcal{P} \nabla^2 T' \label{EQ:linear3}
\end{align}
with Dirichlet boundary conditions 
\begin{equation}
    T'|_{z = \pm \frac{1}{2}} = 0, \quad u'|_{z = \pm \frac{1}{2}} = 0, \quad p'|_{z = \pm \frac{1}{2}} = 0.
\end{equation}
This is now a linear problem in $\vec{u}'$ and $T'$ which can be solved as an eigenvalue problem.

In his groundbreaking report \cite{Rayleigh_1916}, Lord Rayleigh observed that \eqss{EQ:linear1}{EQ:linear3} can be manipulated into a separable form with generalized solutions
\begin{align}
    w'(x, z, t) &= A\, \Re\left[W(z) \, e^{i(k_xx-st)}\right], \label{EQ:normal_modes1}\\ 
    u'(x, z, t) &= A\, \Re\left[U(z) \, e^{i(k_xx-st)}\right], \label{EQ:normal_modes2}\\ 
    T'(x, z, t) &= A\, \Re\left[\theta(z) \, e^{i(k_xx-st)}\right], \label{EQ:normal_modes3}\\ 
    p'(x, z, t) &= A\, \Re\left[P(z) \, e^{i(k_xx-st)}\right], \label{EQ:normal_modes4}
\end{align}
where $A$ is the (undetermined) mode amplitude, $s = \omega + i\sigma$ and $k_x$ is constrained, by periodicity, to the countably infinite set (spectrum) of wavenumbers
\begin{align}
    k_x \in \left\{\frac{n\pi}{2} \, \big| \, n \in \mathbb{N}\right\}.
\end{align}
We normalize the eigenmodes to have
\begin{equation}
  \langle |\theta|^2 \rangle_{\mathcal{D}} = 1.
\end{equation}
For each $k_x$, we can assess the stability of the perturbations by solving for the eigenvalue $s$, whose imaginary component $\sigma$ plays the role of an exponential growth rate. 
Positive eigenvalues indicate that the system is unstable to small disturbances of wavenumber $k_x$, while negative eigenvalues indicate stability. 
A complete linear stability analysis requires solution over the full spectrum of wavenumbers. 
The prototypical case is used to demonstrate that the critical Rayleigh number $\Ra_c = 1708$ when $\frac{\partial \bar{T}}{\partial z} = -1$.

@@

\section{Perturbation Evolution}\label{sec:evolution}
The linearized system \eqss{EQ:linear1}{EQ:linear3} cannot determine the amplitude of the eigenmodes, $A$.
However, the advective heat flux is proportional to $A^2$, so we need to specify the amplitude in order to solve equation~\ref{EQ:T0_IVP}.
To evolve $\bar{T}$, we assume the perturbations evolve on a much faster timescale than the mean temperature, as in \cite{michel_chini_2019}.
Stable modes decay away rapidly. 
Unstable modes will not persist on the slow timescale because the advective term $\langle w'T' \rangle_x$ tends to stabilize $\bar{T}$, thereby creating a negative feedback loop.
Only marginally stable modes can be maintained on the slow timescale.
Therefore the amplitude $A$ must satisfy
\begin{equation}
    \max_{k_x} \{ \sigma \} = 0.
\end{equation}

For various $\Ra$ and fixed $\Pr = 1$, we seek marginally-stable thermal equilibria (MSTE) satisfying $\frac{\partial \bar{T}}{\partial t} = 0$ according to . 
We employ the \texttt{Dedalus} pseudo-spectral python framework \cite{Dedalus_2020} to solve the EVP outlined in Section \ref{sec:model} as well as the IVP \eq{EQ:T0_IVP} by representing each field with Chebyshev polynomials.
We use the 3/2 dealiasing rule to calculate the advective heat flux.
The necessary number of basis functions varies, as eigenfunctions and temperature profiles associated with large $\Ra$ admit increasingly small-scale features. 
We use the \texttt{Eigentools} package \cite{Eigentools} to manipulate the eigenfunctions and calculate the advective heat flux $\langle w' T' \rangle_x$.

At $t = 0$, we begin with a marginally-stable initial profile $\bar{T}(z, 0)$ whose construction is outlined in appendix~\ref{sec:initial_profile}. 
At some arbitrary time $t = t_0$, we seek to evolve $\bar{T}(z, t_0)$ into a new marginally-stable profile $\bar{T}(z, t_0 + \Delta t)$ according to \eq{EQ:T0_IVP}, using backward Euler to evolve the diffusion term and forward Euler to evolve the advection term. 
This requires us to determine the amplitude of the eigenfunction $A$.
We will now illustrate our method of finding $A^2$ through an example.

@@

An iteration is performed as follows.
Consider a marginally-stable temperature profile $\bar{T}(z, t_0)$.
By definition, its maximum eigenvalue is 0.
Diffusing $\bar{T}(z, t_0)$ tends to increase its eigenvalues.
Ignoring the diffusive term and evolving according to advection tends the stabilize the system.
The appropriate amplitude $A^2$ can then be approximated by
\begin{equation}
    A^2 \approx -\frac{\sigma_{\rm{diff}}}{\sigma_{\rm{adv}}} \label{EQ:amp_approx}
\end{equation}
where $\sigma_{\rm{diff}}$ and $\sigma_{\rm{adv}}$ refer to the diffused and advected eigenvalues of the initially marginal mode.
We illustrate these trends in Figure~\ref{fig:iteration_spectra}. 
$\bar{T}(z, t)$ is then evolved according to \eq{EQ:T0_IVP} and another eigenvalue solve is performed. 
Given a fixed timestep $\Delta t$, we assume the dominant eigenvalue can be described by a continuous function $\sigma_{\rm{max}} (A^2)$ which is locally differentiable. 
We use Newton's method to find an amplitude which satisfies our marginal stability tolerance criterion $|\sigma_{\rm{max}} (A^2)| < 10^{-9}$.
Marginally-stable modes do not osscillate in time, i.e. $\sigma = 0$ implies $\omega = 0$.
This agrees with the convectional notion of exchange of stabilities \cite{drazin_reid_2004}.
Crucially, we do not assume the $k_x$ of the marginally-stable is fixed.
In section \ref{sec:multiple_modes} we specify procedures for the treatment of multiple simultaneously marginal modes.

\subsection{Treatment of multiple marginally-stable modes} \label{sec:multiple_modes}
In most cases, we eventually encounter eigenvalue spectra with multiple simultaneously marginal modes.
If ignored, we are forced to reduce the timestep and allow the modes to alternate which is marginally stable.
Instead, we generalize the advective term in \eq{EQ:T0_IVP} to accommodate $N$ simultaneously marginal modes
\begin{equation}
    \langle w' T' \rangle_x = \sum_{n = 1}^{N} 4 A_n^2 \int \, \Re\left[ W_n \theta_n^* \right] \,dz
\end{equation}
where $W_n$ and $\theta_n$ are the eigenmodes associated with $k_x = \frac{n\pi}{2}$. 
There are now $N$ amplitudes to solve for and $N$ eigenvalues to keep marginally-stable. 
Given a small fixed time step $\Delta t$, let $\vec{A^2}$ be the amplitude vector and $\vec{\sigma}(\vec{A^2})$ be the dependent eigenvalues.
We expect a function $\vec{\sigma} \, : \, \mathbb{R}^N \to  \mathbb{R}^N$ to have isolated roots $\vec{A^2_m}$ (should they exist). 
A coarse approximation $\vec{A^2_0}$ for the appropriate amplitude vector $\vec{A^2_m}$ is given by a generalized form of \eq{EQ:amp_approx}:
\begin{equation}
    \vec{A^2_0} = -\Sigma_{\rm{adv}}^{-1} \vec{\sigma_{\rm{diff}}}.
    \label{EQ:AN_approx}
\end{equation}
Here $\vec{\sigma_{\rm{diff}}} = \vec{\sigma} (\vec{0})$ refers to the eigenvalues after a brief $\Delta t$ period of diffusion. 
$\Sigma_{\rm{adv}} \in \mathbb{R}^{N \times N}$ is the matrix of eigenvalues after $N$ periods of advection (one for each mode).
% The $k$th row of $J_0$ is then given by the gradient of the $k$th mode's eigenvalue $\nabla \sigma_k$.
In this way, an arbitrary element of $\Sigma_{\rm{adv}}$ represents how a particular mode's eigenvalue responds to the advection of another mode. 
At the point of our coarse estimate $\vec{A^2_0}$, we approximate the Jacobian matrix 
\begin{equation}
    J = \begin{bmatrix}
        \nabla \sigma_1 (A_1, A_2, ..., A_N) \\
        \nabla \sigma_2 (A_1, A_2, ..., A_N) \\
        \vdots \\
        \nabla \sigma_N (A_1, A_2, ..., A_N) 
    \end{bmatrix}
\end{equation}
via first-order finite differences, requiring $N^2$ more individual EVP solves.
A refined estimate is then given by
\begin{equation}
    \vec{A^2_1} = -J^{-1} \vec{\sigma}(\vec{A^2_0}).
\end{equation}
Once constructed, we use each subsequent guess to adjust the Jacobian via Broyden's method for root-finding in multi-dimensional functions \cite{Broyden}.
We find $\vec{\sigma}(\vec{A^2})$ does indeed have a unique root provided the time step is not too large and there are no numerical instabilities, as outlined in appendix \ref{sec:timestep}.
Presumably this is due to the coupling of $\langle w'T' \rangle_x$ with $\bar{T}$.
Over the course of a large time step, $\bar{T}$ evolves according to \eq{EQ:T0_IVP} and eventually the original eigenfunctions cease to provide a stabilizing influence.

Difficulty arises when transitioning between different numbers of marginal modes, particularly when stable modes become marginally-stable mid-iteration.
We facilitate these transitions by defining an adjustable candidate tolerance $\varepsilon_{\rm{cand}} \in [10^{-6}, 10^{-8}]$.
We rely on $A^2 > 0$ by asserting that a mode which meets the candidate tolerance can be included in an iteration provided that its respective amplitude is positive.
Should that be the case, the candidate mode is now marginally-stable and included in the subsequent iteration.
If however we converge on a negative amplitude, the candidate mode is discarded and the iteration is repeated. 
%In general we find that no other obvious course of action yields equilibria.

After several time steps, $\bar{T}$ tends to asymmetrize due to numerical noise. 
Though asymmetric pairs of solutions may exist, we further constrain $\bar{T}$ by imposing symmetry. 
This is accomplished at the start of each iteration by setting the coefficients of even Chebyshev polynomials to zero.

\section{Properties of Thermally Equilibrated States}\label{sec:properties}
We evolve $\bar{T}$ as described above until $|\partial_{t}\bar{T}| < 10^{-5}$.
In this marginally-stable thermal equilibrium, $\bar{T}$ does not evolve in time, and the perturbations also do not evolve in time, as they are marginally stable.
Thus, these configurations are exact solutions to the quasilinear equations (equations~\ref{EQ:T0_IVP}--\ref{EQ:linear3}).
They differ from the usual ECS in that ECS are fixed points of the full nonlinear problem \eqss{EQ:motion1}{EQ:motion3}. 
Such definitions are not mutually exclusive, but in general we can assume that MSTE and ECS are not steady with respect to their counterparts' equations.
We compute symmetric MSTE for $\Ra$ in the range $10^5 - 10^9$.

@@

Figure \ref{fig:T0_profiles} gives temperature profiles for $\Ra = 10^8$ where the initial profile, whose construction is outlined in appendix~\ref{sec:initial_profile}, is employed at iteration 0. 
Direct numerical simulations (DNS) are performed by solving \eqss{EQ:motion1}{EQ:motion3} with \texttt{Dedalus}, followed by horizontal- and time- averaging. 
The DNS curve is more diffuse than the MSTE curve, which in turn, is more diffuse than the initial curve. 
Performing an eigenvalue solve by setting $\bar{T}$ equal to the DNS profile yields unstable eigenvalues. 
This suggests that MSTE might maximize boundary layer thickness, subject to the marginal stability constraint.

>>>>

\author{Liam O'Connor$^1$}
\author{Daniel Lecoanet$^{1, 2}$}
\author{Evan H. Anders$^2$}
\affiliation{%
$^1$Department of Engineering Sciences and Applied Mathematics, Northwestern University, Evanston, IL 60208 USA}
\affiliation{%
$^2$Center for Interdisciplinary Exploration and Research in Astrophysics, Northwestern University, Evanston, IL, 60201 USA}

\begin{abstract}
    Natural convection is ubiquitous throughout the physical sciences and engineering, yet many of its important properties remain elusive---particularly in the large Rayleigh number regime.
    In this investigation, we derive and solve a quasilinear form of the Rayleigh-Bénard problem by representing the perturbations in terms of marginally stable eigenmodes.
    The amplitude of each eigenmode is determined by requiring that the background state maintains marginal stability.
    The background temperature profile evolves due to the advective flux of each eigenmode, as well as diffusion.
    The entire calculation is one-dimensional, and can be run on a workstation.
    We find the background temperature field evolves to an equilibrium state, where the advective flux from the marginally-stable eigenmodes and the diffusive flux sum to a constant.
    These marginally-stable thermal equilibria are exact solutions of the quasilinear equations.
    The mean temperature profile has thinner boundary layers and larger Nusselt numbers than thermally-equilibrated 2D and 3D simulations of the full nonlinear equations.
    We find the Nusselt number scales like $\Nu \sim\Ra^{1/3}$.
    When initializing a 2D simulation with a marginally-stable thermal equilibrium, we find the simulation does not exhibit the initial burst of turbulence typical of simulations initialized with the conductive background state plus noise.
    Although the $\Nu$ quickly equilibrates in these simulations, the kinetic energy evolves on a viscous timescale as domain-filling flywheel flows viscously attenuate.
\end{abstract}

@@

\section{Introduction}
Rayleigh-B\'enard convection plays a foundational role in astrophysical and geophysical settings.
The resulting buoyancy-driven flows regulate heat transfer and generate large-scale vortices \cite{Couston}.
Turbulent convection, which is associated with large Rayleigh numbers $\Ra$, is difficult to simulate. 
State of the art simulations performed by \cite{Zhu_2018} have reached $\Ra \sim 10^{14}$ but estimates for the sun's convective zone and earth's interior are $\Ra \sim 10^{16}-10^{20}$ and $\Ra \sim 10^{20}-10^{30}$ respectively \cite{Ossendrijver,Gubbins_2001}. 
The scaling behavior of the Nusselt number $\Nu \sim \Ra^{\beta}$ in the asymptotic ultimate regime is of particular interest.
There exists a substantial body of work pertaining to this specific topic with no general consensus \cite{Malkus_1954, Howard_1966, Kraichnan, Spiegel, Castaing, Grossman, Ahlers}.

@@

\section{Model Setup}\label{sec:model}
We begin with the Boussinesq approximation for Rayleigh-Bénard Convection, nondimensionalized on the freefall timescale. 
The domain $\mathcal{D}$ is 2-dimensional, rectangular, and horizontally periodic with spatial dimensions $0 < x < 4$ and $-1/2 < z < 1/2$. 
The fluid of interest is constrained between two flat boundaries at $z = -1/2$ and $z = 1/2$ with fixed temperatures $1/2$ and $-1/2$ respectively. 
At both boundaries we specify impenetrable, no-slip conditions, such that the velocity $\vec{u} = u \hat{x} + w \hat{z} = \vec{0}$ at $z = \pm 1/2$, where $\hat{x}, \hat{z}$ are the unit vectors in the $x$ and $z$ directions. 
The equations of motion are then given by
\begin{align}
    \nabla \cdot \vec{u} &= 0 \label{EQ:motion1}\\
    \frac{\partial \vec{u}}{\partial t} + \vec{u} \cdot \nabla \vec{u} &= - \nabla p + T \hat{z} + \mathcal{R} \nabla^2 \vec{u} \label{EQ:motion2}\\
    \frac{\partial T}{\partial t} + \vec{u} \cdot \nabla T &= \mathcal{P} \nabla^2 T \label{EQ:motion3}
\end{align}
where $p$ is pressure and $T$ is temperature. 
For completeness, we specify a final boundary condition $p = p_0$ at $z = \pm 1/2$. 
Any system of this form can be characterized by its dimensionless Rayleigh number $\Ra = \frac{g\alpha L^3 \Delta T}{\nu \kappa}$ and Prandtl number $\Pr = \frac{\nu}{\kappa}$, where $g, \, \alpha, \, L, \, \Delta T, \nu, \kappa$ are the gravitational acceleration, coefficient of thermal expansion, domain height, opposed temperature difference, kinematic viscosity, and thermal diffusivity respectively. 
In this paper, we fix $\Pr = 1$.
For convenience, we define
\begin{equation}
\mathcal{R} = \sqrt{\frac{\Pr}{\Ra}}, \qquad \mathcal{P} = \frac{1}{\sqrt{\Pr \Ra}}.
\end{equation}

@@

Substituting \eq{EQ:reynolds_dc_p} into \eq{EQ:motion2} and taking the horizontal average reveals that the mean pressure field $\bar{p}(z)$ must satisfy
\begin{equation}
    0 = -\frac{\partial \bar{p}}{\partial z} + \bar{T}\label{EQ:p_bar}
\end{equation}

To solve \eq{EQ:T0_IVP} numerically, we need an expression for the perturbation so we can calculate the advective heat flux.
Here we will make the quasilinear approximation, dropping the $\vec{u}'\vec{\cdot}\vec{\nabla}\vec{u}'$ and $\vec{u}'\vec{\cdot}\vec{\nabla}T'$ terms from the evolution equations for the perturbations.
Substituting \eqss{EQ:reynolds_dc_u}{EQ:reynolds_dc_p} into \eqss{EQ:motion1}{EQ:motion3} followed by subtracting \eq{EQ:T0_IVP} and \eq{EQ:p_bar} from the resulting temperature and $\hat{z}$ momentum equations gives
\begin{align}
    \nabla \cdot \vec{u'} &= 0 \label{EQ:linear1}\\
    \frac{\partial\vec{u'}}{\partial t} &= - \nabla p' + T'\hat{z} + \mathcal{R} \nabla^2 \vec{u'} \label{EQ:linear2}\\
    \frac{\partial T'}{\partial t} + \frac{\partial \bar{T}}{\partial z} w' &= \mathcal{P} \nabla^2 T' \label{EQ:linear3}
\end{align}
with Dirichlet boundary conditions 
\begin{equation}
    T'|_{z = \pm \frac{1}{2}} = 0, \quad u'|_{z = \pm \frac{1}{2}} = 0, \quad p'|_{z = \pm \frac{1}{2}} = 0.
\end{equation}
This is now a linear problem in $\vec{u}'$ and $T'$ which can be solved as an eigenvalue problem.

In his groundbreaking report \cite{Rayleigh_1916}, Lord Rayleigh observed that \eqss{EQ:linear1}{EQ:linear3} can be manipulated into a separable form with generalized solutions
\begin{align}
    w'(x, z, t) &= A\, \Re\left[W(z) \, e^{i(k_xx-st)}\right], \label{EQ:normal_modes1}\\ 
    u'(x, z, t) &= A\, \Re\left[U(z) \, e^{i(k_xx-st)}\right], \label{EQ:normal_modes2}\\ 
    T'(x, z, t) &= A\, \Re\left[\theta(z) \, e^{i(k_xx-st)}\right], \label{EQ:normal_modes3}\\ 
    p'(x, z, t) &= A\, \Re\left[P(z) \, e^{i(k_xx-st)}\right], \label{EQ:normal_modes4}
\end{align}
where $A$ is the (undetermined) mode amplitude, $s = \omega + i\sigma$ and $k_x$ is constrained, by periodicity, to the countably infinite set (spectrum) of wavenumbers
\begin{align}
    k_x \in \left\{\frac{n\pi}{2} \, \big| \, n \in \mathbb{N}\right\}.
\end{align}
We normalize the eigenmodes to have
\begin{equation}
  \langle |\theta|^2 \rangle_{\mathcal{D}} = 1
\end{equation}
where $\langle \cdot \rangle_{\mathcal{D}}$ denotes the spatial mean over the entire domain.

For each $k_x$, we can assess the stability of the perturbations by solving for the eigenvalue $s$, whose imaginary component $\sigma$ plays the role of an exponential growth rate. 
Positive eigenvalues indicate that the system is unstable to small disturbances of wavenumber $k_x$, while negative eigenvalues indicate stability. 
A complete linear stability analysis requires solution over the full spectrum of wavenumbers. 
The prototypical case is used to demonstrate that the critical Rayleigh number $\Ra_c = 1708$ when $\frac{\partial \bar{T}}{\partial z} = -1$.

@@

\section{Perturbation Evolution}\label{sec:evolution}
The linearized system \eqss{EQ:linear1}{EQ:linear3} cannot determine the amplitude of the eigenmodes, $A$.
However, the advective heat flux is proportional to $A^2$, so we need to specify the amplitude in order to solve equation~\ref{EQ:T0_IVP}.
To evolve $\bar{T}$, we assume the perturbations evolve on a much faster timescale than the mean temperature, as in \cite{michel_chini_2019}.
Stable modes $(\sigma < 0)$ decay away rapidly. 
Unstable modes $(\sigma > 0)$ will not persist on the slow timescale because the advective term $\langle w'T' \rangle_x$ tends to stabilize $\bar{T}$, thereby creating a negative feedback loop.
Only marginally stable modes can be maintained on the slow timescale.
Therefore the amplitude $A$ must satisfy
\begin{equation}
    \max_{k_x} \{ \sigma \} = 0.
\end{equation}

For various $\Ra$ and fixed $\Pr = 1$, we seek marginally-stable thermal equilibria (MSTE) satisfying $\frac{\partial \bar{T}}{\partial t} = 0$ according to \eq{EQ:T0_IVP}. 
We employ the \texttt{Dedalus} pseudo-spectral python framework \cite{Dedalus_2020} to solve the EVP outlined in Section \ref{sec:model} as well as the IVP \eq{EQ:T0_IVP} by representing each field with Chebyshev polynomials.
We use the 3/2 dealiasing rule to calculate the advective heat flux.
The necessary number of basis functions varies, as eigenfunctions and temperature profiles associated with large $\Ra$ admit increasingly small-scale features. 
A two-stage Runge-Kutta timestepper is implemented for the IVP.
We use the \texttt{Eigentools} package \cite{Eigentools} to manipulate the eigenfunctions and calculate the advective heat flux $\langle w' T' \rangle_x$.

At $t = 0$, we begin with a marginally-stable initial profile $\bar{T}(z, 0)$ whose construction is outlined in appendix~\ref{sec:initial_profile}. 
At some arbitrary time $t = t_0 \geq 0$, we seek to evolve $\bar{T}(z, t_0)$ into a new marginally-stable profile $\bar{T}(z, t_0 + \Delta t)$ according to \eq{EQ:T0_IVP}, using backward Euler to evolve the diffusion term and forward Euler to evolve the advection term. 
This requires us to determine the amplitude of the eigenfunction $A$.
We will now illustrate our method of finding $A^2$ through an example.

@@

An iteration is performed as follows.
Consider a marginally-stable temperature profile $\bar{T}(z, t_0)$.
By definition, its maximum eigenvalue is 0.
Diffusing $\bar{T}(z, t_0)$ tends to increase its eigenvalues.
Ignoring the diffusive term and evolving according to advection tends the stabilize the system.
The appropriate amplitude $A^2$ can then be approximated by
\begin{equation}
    A^2 \approx -\frac{\sigma_{\rm{diff}}}{\sigma_{\rm{adv}}} \label{EQ:amp_approx}
\end{equation}
where $\sigma_{\rm{diff}}$ and $\sigma_{\rm{adv}}$ refer to the diffused and advected eigenvalues of the initially marginal mode.
We illustrate these trends in Figure~\ref{fig:iteration_spectra}. 
$\bar{T}(z, t)$ is then evolved according to \eq{EQ:T0_IVP} and another eigenvalue solve is performed. 
Given a fixed timestep $\Delta t$, we assume the dominant eigenvalue can be described by a continuous function $\sigma_{\rm{max}} (A^2)$ which is locally differentiable. 
We use Newton's method to find an amplitude which satisfies our marginal stability tolerance criterion $|\sigma_{\rm{max}} (A^2)| < 10^{-9}$.
Marginally-stable modes do not oscillate in time, i.e. $\sigma = 0$ implies $\omega = 0$.
This agrees with the conventional notion of exchange of stabilities \cite{drazin_reid_2004}.
Crucially, we do not assume the $k_x$ of the marginally-stable mode is fixed.
In section \ref{sec:multiple_modes} we specify procedures for the treatment of multiple simultaneously marginal modes.

\subsection{Treatment of multiple marginally-stable modes} \label{sec:multiple_modes}
In most cases, we encounter eigenvalue spectra with multiple simultaneously marginal modes.
If ignored, we are forced to reduce the timestep and allow the modes to alternate which is marginally stable.
Instead, we generalize the advective term in \eq{EQ:T0_IVP} to accommodate $N$ simultaneously marginal modes
\begin{equation}
    \langle w' T' \rangle_x = \sum_{n = 1}^{N} 4 A_n^2 \int \, \Re\left[ W_n \theta_n^* \right] \,dz
\end{equation}
where $W_n$ and $\theta_n$ are the eigenmodes associated with $k_x = \frac{n\pi}{2}$. 
There are now $N$ amplitudes to solve for and $N$ eigenvalues to keep marginally-stable. 
Given a small fixed time step $\Delta t$, let $\vec{A^2}$ be the amplitude vector and $\vec{\sigma}(\vec{A^2})$ be the dependent eigenvalues.
We expect a function $\vec{\sigma} \, : \, \mathbb{R}^N \to  \mathbb{R}^N$ to have isolated roots $\vec{\tilde{A}^2}$ (should they exist). 
A coarse approximation $\vec{A^2_0}$ for the appropriate amplitude vector $\vec{\tilde{A}^2}$ is given by a generalized form of \eq{EQ:amp_approx}:
\begin{equation}
    \vec{\tilde{A}^2} \approx \vec{A^2_0} = -\Sigma_{\rm{adv}}^{-1} \vec{\sigma_{\rm{diff}}}.
    \label{EQ:AN_approx}
\end{equation}
Here $\vec{\sigma_{\rm{diff}}} = \vec{\sigma} (\vec{0})$ refers to the eigenvalues after a brief $\Delta t$ period of diffusion. 
$\Sigma_{\rm{adv}} \in \mathbb{R}^{N \times N}$ is the matrix of eigenvalues after $N$ periods of advection (one for each mode).
% The $k$th row of $J_0$ is then given by the gradient of the $k$th mode's eigenvalue $\nabla \sigma_k$.
In this way, an arbitrary element of $\Sigma_{\rm{adv}}$ represents how a particular mode's eigenvalue responds to the advection of another mode. 
At the point of our coarse estimate $\vec{A^2_0}$, we approximate the Jacobian matrix 
\begin{equation}
    J = \begin{bmatrix}
        \nabla \sigma_1 (A_1, A_2, ..., A_N) \\
        \nabla \sigma_2 (A_1, A_2, ..., A_N) \\
        \vdots \\
        \nabla \sigma_N (A_1, A_2, ..., A_N) 
    \end{bmatrix}
\end{equation}
via first-order finite differences, requiring $N^2$ more individual EVP solves.
A refined estimate is then given by
\begin{equation}
    \vec{\tilde{A}^2} \approx \vec{A^2_1} = -J^{-1} \vec{\sigma}(\vec{A^2_0}).
\end{equation}
Once constructed, we use each subsequent guess to adjust the Jacobian via Broyden's method for root-finding in multi-dimensional functions \cite{Broyden}.
We find $\vec{\sigma}(\vec{A^2})$ does indeed have a unique root provided the time step is not too large and there are no numerical instabilities, as outlined in appendix \ref{sec:timestep}.
Presumably this is due to the coupling of $\langle w'T' \rangle_x$ with $\bar{T}$.
Over the course of a large time step, $\bar{T}$ evolves according to \eq{EQ:T0_IVP} and eventually the original eigenfunctions cease to provide a stabilizing influence.

Difficulty arises when transitioning between different numbers of marginal modes, particularly when stable modes become marginally-stable mid-iteration.
We facilitate these transitions by defining an adjustable candidate tolerance $\varepsilon_{\rm{cand}} \in [10^{-6}, 10^{-8}]$.
We rely on $A^2 > 0$ by asserting that a mode which meets the candidate tolerance can be included as a candidate marginal mode.
If the candidate mode's amplitude turns out to be positive, the candidate mode is now marginally-stable and can be regularly included in the subsequent iteration.
If however we converge on a negative amplitude, the candidate mode must be discarded and the iteration is repeated.

After several time steps, $\bar{T}$ tends to asymmetrize due to numerical noise. 
Though asymmetric pairs of solutions may exist, we further constrain $\bar{T}$ by setting the coefficients of its even Chebyshev basis functions to zero at the start of each iteration.

\section{Properties of Thermally Equilibrated States}\label{sec:properties}
We evolve $\bar{T}$ as described above until $\max\{|\partial_{t}\bar{T}|\} < 10^{-5}$.
In this marginally-stable thermal equilibrium, $\bar{T}$ does not evolve in time, and the perturbations also do not evolve in time, as they are marginally stable.
Thus, these configurations are exact solutions to the quasilinear equations (equations~\ref{EQ:T0_IVP}--\ref{EQ:linear3}).
They differ from the usual ECS in that ECS are fixed points of the full nonlinear problem \eqss{EQ:motion1}{EQ:motion3}. 
Such definitions are not mutually exclusive, but in general we can assume that MSTE and ECS are not steady with respect to their counterparts' equations.
We compute symmetric MSTE for $\Ra$ in the range $10^5 - 10^9$.

@@

Figure~\ref{fig:T0_profiles} gives temperature profiles for $\Ra = 10^8$ where the initial profile, whose construction is outlined in appendix~\ref{sec:initial_profile}, is employed at iteration 0. 
Direct numerical simulations (DNS) are performed by solving \eqss{EQ:motion1}{EQ:motion3} with \texttt{Dedalus}, followed by horizontal- and time- averaging. 
The DNS curve is more diffuse than the MSTE curve, which in turn, is more diffuse than the initial curve. 
Performing an eigenvalue solve by setting $\bar{T}$ equal to the DNS profile yields unstable eigenvalues. 
This suggests that MSTE might maximize boundary layer thickness, subject to the marginal stability constraint.
